---
title: "Explore Data"
author: "Aleksei Sorokin, asorokin@hawk.iit.edu, A20394300"
date: "3/4/2020"
output: pdf_document
geometry: margin=1in
---

## Load data and do some final filtering
```{r}
df <- read.csv('data/all_coaches.csv')
data_dict <- list(
              N='Name',
              GR='Games Relative',
              WP='Win Loss Percentage',
              PGR='Playoff Games Relative',
              PWP='Playoff Win Percentage',
              CC='Conference Championships',
              C='Championships',
              HOF='Hall of Fame',
              S='Sport')
names(df) <- names(data_dict)
# export stats
head(df)
nrow(df) # total coaches
sum(df$HOF==1) # total hall of fame coaches
sum(df$HOF==0) # total non hall of fame coaches
summary(df)
```

## Correlation between data
```{r}
library(corrplot)
df_num <- Filter(is.numeric,df)
df_corr <- cor(df_num)
corrplot(df_corr)
```

## Boxplots over Histograms
```{r}
fixed_cols <- names(df_num)[names(df_num)!='HOF']
for (col in fixed_cols){
  # gather data
  c_name <- data_dict[[col]]
  data <- df_num[[col]]
  # make plot
  layout(mat = matrix(c(1,2),2,1, byrow=TRUE),  height = c(1,8))
  par(mar=c(0, 3.1, 1.1, 2.1))
  boxplot(data , horizontal=TRUE , xaxt="n", frame=F,main=sprintf('Histogram of %s',c_name))
  par(mar=c(4, 3.1, 1.1, 2.1))
  hist(data,xlab=c_name,main='')
  # print top coaches in this category
  cat(sprintf('Top 10 coaches by %s\n',c_name))
  df_top <- df[order(df[[col]],decreasing=T),] 
  print(df_top[1:10,])}
```

## Some potential problems
- Dataset does not capture coach's personalities, effect on players, effect on the league, work in community, ...
- Many coaches are in HOF despite having less than a full year of head coaching.This is probably due to their assistant coaching experience which is not captured in this dataset
  - It would be nice to filter this dataset to coaches who have only over a year experience to account for outlier such as coaches with a win-loss percentage of 1 because they have only coached one game. However, this would remove almost 50 coaches who made the hall f fame despite having less than a year of head coaching experience
```{r}
head(df[df$GR<1 & df$HOF==1,])
```
